Graph-Based Neural Reconstruction from Skeletonized 3D Networks

### 2018-03-12 papers/eccv2018/2_method.tex
papers: major updates to methods

The simplest node generation strategy creates one node for every unique segment label in the input volume. 
However, some of the labels in the volume correspond to very small structures that are likely the result of segmentation errors, typically in regions with noisy raw image data. 
It is difficult to extract useful shape features from these segments because of their small, often random, shape. 
We prune these nodes from the graph by removing all segments with fewer than a threshold $t_{seg} = 20,000$ voxels (\TODO{~XX\SI{}{\nano\meter^3}}). 
This removed on average \TODO{XX}\% of the segments in our  datasets (Sec.~\ref{sec:dataset}). Despite the large number of segments, these regions only take up \TODO{XX}\% of the total volume on average.
\TODO{How much does this change VI - should that matter?}
\TODO{Picture of these examples}

@@

A typical approach for generating edges produces one between all adjacent segments. Two segments $l_1$ and $l_2$ are considered adjacent if there is a pair of adjacent voxels with one labeled $l_1$ and the other labeled $l_2$.
For example, pixel-based agglomeration methods such as NeuroProof~\cite{10.1371/journal.pone.0125825} and GALA~\cite{nunez2014graph} consider all pairs of adjacent segments for merging.
However, this method produces too many edges in the graph for graph-based optimization approaches. 
We identify a smaller number of pairs of segments to consider as graph edges using the following approach.
\TODO{address comment from that one reviewer}

First, we extract a skeleton from each segment in the label volume using a variant~\cite{zhao2014automatic} of the TEASER algorithm~\cite{sato2000teasar}. 
This skeletonization algorithm repeatedly uses Dijkstra's algorithm to find the farthest voxel from a seed location. 
Since this algorithm is non-linear in the number of voxels, we downsample the datasets so that there are voxel samples every $\SI{30}{\nano\meter}$ in each dimension.
Empirically, this reduced the running time for skeletonization by $\TODO{XX\%}$, with minimal reduction in skeleton accuracy ($\TODO{XX\%}$ fewer branches). 
Fig.~\ref{fig:skeletonization} shows three examples of extracted skeletons (in black). 
These skeletons consist of a sequence of \textit{joints}, i.e., locations that are locally a maximum distance from the segment boundary, with line segments connecting successive joints. 
We prune the joints that are within $t_{jnt} = \TODO{XX}\SI{}{\nano\meter}$ of each other to reduce unnecessary branching. 
\TODO{why this value?}
We refer to joints that have only one connected neighbor as \textit{endpoints}. 
Many of the segments that are erroneously split have nearby endpoints  (Fig.~\ref{fig:merge_candidates}). 
\TODO{Numbers for this?}
We make use of this fact to find merge candidates with the following two-pass pruning algorithm.

In the first pass, we iterate over all endpoints $e$ belonging to a segment $S$ and create a set of segments $\mathbb{S}_e^\prime$ that includes all labels that are within $t_{low}$ voxels from $e$.
Elements of $\mathbb{S}_e^\prime$ are candidates for merging. 
However, this first pass often leads to too many candidates, requiring an additional pass for further pruning. 
In the second pass, we consider all of the segments in $\mathbb{S}_e^\prime$ for every endpoint $e$. 
If a segment $S^\prime \in \mathbb{S}_e^\prime$ has an endpoint within $t_{high}$ voxels of $e$, the segment $S$ and $S^\prime$ are considered for merging. 
We store the midpoints between the two endpoints as the center of the potential merge in the set $\mathbb{S}_c$. 
This algorithm produces a set of segments to consider for merging. Only these pairs have a corresponding edge in the constructed graph.
\TODO{From Daniel: this sounds like a double threshold to me.. imho 1-2 sentences without formulas is clearer?}

\subsection{Edge Weights Assignment}
We assign edge weights $w_e$ to each edge where the weight corresponds to the probability that two nodes belong to the same neuron.
We train a 3D CNN classifier to learn from the manually labeled oversegmentation input volume (Sec.~\ref{sec:dataset}).
If the probability that the nodes belong to the same neuron is $p_e$, the edge weight $w_e = \log{\frac{p_e}{1 - p_e}} + \log{\frac{1 - \beta}{\beta}}$, where $\beta$ is a tunable parameter that encourages over- or under-segmentation.

@@

\subsection{Graph Partitioning}

@@

By reformulating the segmentation problem as a graph partitioning one, we can enforce some global constraints on our result based on the underlying biology.
Traditional hierarchical clustering algorithms do not rely on such constraints but consider local decisions independently.
The multicut problems returns a series of ``collapsed'' edges between nodes that belong to the same neuron.
We iterate over these edges in order of the probability of merge generated by our CNN. 
\TODO{HOW DO WE ACCOMPLISH THIS}

>>>>

The simplest node generation strategy creates one node for every unique segment label in the input volume. 
However, some of the labels in the volume correspond to very small structures that are likely the result of segmentation errors, typically in regions with noisy raw image data. 
It is difficult to extract useful shape features from these segments because of their small, often random, shape. 
We prune these nodes from the graph by removing all segments with fewer than a threshold $t_{seg} = 20,000$ voxels. 
This removes on average \TODO{XX}\% of the segments in our  datasets (Sec.~\ref{sec:dataset}), leaving us with around $\TODO{XX}$ nodes per dataset (\TODO{X nodes per micrometer}). 
Despite the large number of segments, these regions only take up \TODO{XX}\% of the total volume on average.
We provide further analysis of this strategy in the supplemental material.

@@

A typical approach for generating edges produces one between all adjacent segments. Two segments $l_1$ and $l_2$ are considered adjacent if there is a pair of adjacent voxels with one labeled $l_1$ and the other labeled $l_2$.
For example, pixel-based agglomeration methods such as mean agglomeration \TODO{CITE} or waterz \TODO{CITE} consider all pairs of adjacent segments for merging.
However, this method produces too many edges in the graph for graph-based optimization approaches. 
We identify a smaller number of pairs of segments to consider as graph edges using the following approach.

First, we extract a skeleton from each segment in the label volume using a variant~\cite{zhao2014automatic} of the TEASER algorithm~\cite{sato2000teasar}. 
This skeletonization algorithm repeatedly uses Dijkstra's algorithm to find the farthest voxel from a seed location. 
Since this algorithm is non-linear in the number of voxels, we downsample the datasets so that there are voxel samples every $\SI{30}{\nano\meter}$ in each dimension.
Empirically, this reduced the running time for skeletonization by $\TODO{XX\%}$, with minimal reduction in skeleton accuracy ($\TODO{XX\%}$ fewer branches). 
Fig.~\ref{fig:skeletonization} shows three examples of extracted skeletons (in black). 
These skeletons consist of a sequence of \textit{joints}, i.e., locations that are locally a maximum distance from the segment boundary, with line segments connecting successive joints. 
We prune the joints that are within $t_{jnt} = \TODO{XX}\SI{}{\nano\meter}$ of each other to reduce unnecessary branching.
This value is small enough that the skeletons extend into the spines but large enough to keep branching does to a reasonable level.
Additional analysis is available in the supplemental material.
We refer to joints that have only one connected neighbor as \textit{endpoints}. 
We find that approximately \TODO{XX\%} of the segments that are erroneously split have nearby endpoints  (Fig.~\ref{fig:merge_candidates}). 
We make use of this fact to find merge candidates with the following two-pass pruning algorithm.

In the first pass, we iterate over all endpoints $e$ belonging to a segment $S$ and create a set of segments $\mathbb{S}_e^\prime$ that includes all labels that are within $t_{low}$ voxels from $e$.
Elements of $\mathbb{S}_e^\prime$ are candidates for merging. 
However, this first pass often leads to too many candidates, requiring an additional pass for further pruning. 
In the second pass, we consider all of the segments in $\mathbb{S}_e^\prime$ for every endpoint $e$. 
If a segment $S^\prime \in \mathbb{S}_e^\prime$ has an endpoint within $t_{high}$ voxels of $e$, the segment $S$ and $S^\prime$ are considered for merging. 
We store the midpoints between the two endpoints as the center of the potential merge in the set $\mathbb{S}_c$. 
This algorithm produces a set of segments to consider for merging. Only these pairs have a corresponding edge in the constructed graph.

\subsection{Edge Weights Assignment}
\label{sec:edge-weights}
We assign edge weights $w_e$ to each edge where the weight corresponds to the probability that two nodes belong to the same neuron.
We train a 3D CNN classifier to learn from the manually labeled oversegmentation input volume (Sec.~\ref{sec:dataset}).
If the probability that the nodes belong to the same neuron is $p_e$, the edge weight $w_e = \log{\frac{p_e}{1 - p_e}} + \log{\frac{1 - \beta}{\beta}}$, where $\beta$ is a tunable parameter that encourages over- or under-segmentation.

@@

\subsection{Graph Clustering}

@@

By reformulating the segmentation problem as a graph partitioning one, we can enforce some global constraints on our result based on the underlying biology.
Traditional hierarchical clustering algorithms do not rely on such constraints but consider local decisions independently.
We enforce a global constraint that neurons are tree-structured and should not contain cycles. 
The multicut problems returns a series of ``collapsed'' edges between nodes that belong to the same neuron.
We iterate over these edges in order of the probability of merge generated by our CNN. 
We ``collapse'' the edge only if it does not create a cycle in the graph.

### 2018-03-12 papers/eccv2018/3_evaluation.tex
papers: major updates to methods

% !TEX root =  paper.tex
\section{Experimental Results}
\TODO{right now, too many subsections. shall we break it into "Setup" and "Results"}
We evaluate our method by comparing it to a state-of-the-art pixel-based reconstruction approach using datasets from mouse and fly brains.

@@

@@

%The segmentation of the Kasthuri dataset was computed by agglomerating 3D supervoxels produced by the watershed algorithm from 3D affinity predictions~\cite{zlateski2015image}. 
%A recent study by Funke et al.~\cite{funke2017deep} demonstrated superior performance of such methods over existing ones on anisotropic data. 
%We learn 3D affinities using MALIS loss with a U-net~\cite{ronneberger2015u,Turaga:2009}. 
%We apply the z-watershed algorithm with suitable parameters to compute a 3D oversegmentation of the volume. 
%The resulting 3D oversegmentation is then agglomerated using the technique of context-aware delayed agglomeration to generate the final segmentation~\cite{10.1371/journal.pone.0125825}.
For the Kasthuri dataset, we used the segmentation pipeline from Funke et al.~\cite{funke2017deep}. 
We first trained a 3D affinity prediction U-Net model~\cite{ronneberger2015u} with Malis loss~\cite{Turaga:2009} and applied the watershed and agglomeration method~\cite{funke2017deep} to obtain the final segmentation result.

For the FlyEM data, based on the authors' suggestion~\cite{takemura2017connectome}, we applied a context-aware delayed agglomeration algorithm~\cite{10.1371/journal.pone.0125825} that shows improved performance on this dataset over the pipeline used in the original publication. 
This segmentation framework learns voxel and supervoxel classifiers with an emphasis to minimize under-segmentation error. 
The algorithm first computes multi-channel 3-D predictions for membranes, cell interiors, and mitochondria, among other cell features. 
The membrane prediction channel is used to produce an over-segmented volume using 3D watershed, which is then agglomerated hierarchically up to a certain confidence threshold. 
We used exactly the same parameters as the publicly available code for this algorithm.

@@

We use half of the Kasthuri dataset for training and validation. 
We train on 80\% of the potential merge candidates for this volume.
We validate the CNN classifier on the remaining 20\% of candidates.

@@

\TODO{we don't have space for the table for training parameters: move to supp}
\begin{table}[h!]
	\centering
	\begin{tabular}{l l} \hline
		\textbf{Parameters} & \textbf{Values} \\ \hline
		Loss Function & Mean Squared Error \\
		Optimizer & SGD  with Nesterov Momentum \\
		Momentum & 0.9 \\
		Initial Learning Rate & 0.0001 \\
		Decay Rate & $5 * 10^{-8}$ \\
		Activation & LeakyReLU $(\alpha = 0.001)$ \\
		Kernel Sizes & $3 \times 3 \times 3$ \\
		Filter Sizes & $16 \to 32 \to 64$ \\ \hline
	\end{tabular}
	\caption{Training parameters.}
	\label{table:architecture}
\end{table}

@@

>>>>

% !TEX root =  paper.tex
\section{Experiments}

We evaluate our method by comparing it to a state-of-the-art pixel-based reconstruction approach using datasets from mouse and fly brains.

@@

\paragraph{Fib-25.}
The Fib-25 dataset comes from the visual system of a \textit{Drosophila} fly~\cite{takemura2015synaptic}. This sample is imaged at a resolution of $\SI[product-units=single]{8 x 8 x 8}{\nano\meter}^3$ and has $2588 \times 2108 \times 2876$ voxels representing a volume of $\SI[product-units=single]{23.0 x 16.9 x 23.0}{\micro\meter}^3$. 
There is currently a reconstruction challenge based on this sample.

@@

%The segmentation of the Kasthuri dataset was computed by agglomerating 3D supervoxels produced by the watershed algorithm from 3D affinity predictions~\cite{zlateski2015image}. 
%A recent study by Funke et al.~\cite{funke2017deep} demonstrated superior performance of such methods over existing ones on anisotropic data. 
%We learn 3D affinities using MALIS loss with a U-net~\cite{ronneberger2015u,Turaga:2009}. 
%We apply the z-watershed algorithm with suitable parameters to compute a 3D oversegmentation of the volume. 
%The resulting 3D oversegmentation is then agglomerated using the technique of context-aware delayed agglomeration to generate the final segmentation~\cite{10.1371/journal.pone.0125825}.
For the Kasthuri dataset, we used the segmentation pipeline from Funke et al.~\cite{funke2017deep}. 
We first trained a 3D affinity prediction U-Net model~\cite{ronneberger2015u} with MALIS loss~\cite{Turaga:2009} and applied the watershed and agglomeration method~\cite{funke2017deep} to obtain the final segmentation result.
For the FlyEM data, based on the authors' suggestion~\cite{takemura2017connectome}, we applied a context-aware delayed agglomeration algorithm~\cite{10.1371/journal.pone.0125825} that shows improved performance on this dataset over the pipeline used in the original publication. 
This segmentation framework learns voxel and supervoxel classifiers with an emphasis to minimize under-segmentation error. 
The algorithm first computes multi-channel 3-D predictions for membranes, cell interiors, and mitochondria, among other cell features. 
The membrane prediction channel is used to produce an over-segmented volume using 3D watershed, which is then agglomerated hierarchically up to a certain confidence threshold. 
We used exactly the same parameters as the publicly available code for this algorithm.
For the Fib-25 challenge dataset, we use the segmentation from Funke et al. as our input~\cite{funke2017deep}.

@@

We use half of the Kasthuri and FlyEM datasets\footnote{Since the FlyEM datasets have significantly fewer examples, we initialize the network with the weights from the Kasthuri training and have an initial learning rate of $10^{-4}$.} for training and validation. 
We train on 80\% of the potential merge candidates for this volume.
We validate the CNN classifier on the remaining 20\% of candidates.

@@

\begin{table}[h!]
	\centering
	\begin{tabular}{l l} \hline
		\textbf{Parameters} & \textbf{Values} \\ \hline
		Loss Function & Mean Squared Error \\
		Optimizer & SGD  with Nesterov Momentum \\
		Momentum & 0.9 \\
		Initial Learning Rate & 0.0001 \\
		Decay Rate & $5 * 10^{-8}$ \\
		Activation & LeakyReLU $(\alpha = 0.001)$ \\
		Kernel Sizes & $3 \times 3 \times 3$ \\
		Filter Sizes & $16 \to 32 \to 64$ \\ \hline
	\end{tabular}
	\caption{Training parameters.}
	\label{table:architecture}
\end{table}

@@

\subsection{Error Metric}
\label{sec:variation-of-information}

We evaluate the performance of the different methods using the split variation of information (VI)~\cite{meila2003comparing}.
Given a ground truth labeling $GT$ and our automatically reconstructed segmentation $SG$, over- and under-segmentation are quantified by the conditional entropies $H(GT | SG)$ and $H(SG | GT)$, respectively. 
Since we are measuring the entropy between two clusterings, lower VI scores are better.
